---
title: "Perachora Overview Mapping"
author: "Adela Sobotkova"
date: "07/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(raster)
```

## Load Greek background data
```{r load-gr-data}
# Greek border
border <- st_read("data/Greece/periphereies/periphereies.shp", options = "ENCODING=WINDOWS-1253")
border <- st_transform(border, crs = st_crs(features))
border_ppap <- st_crop(border,e)
border_sm <- st_simplify(st_crop(border,e), dTolerance = 150)

# aktog <- st_read("data/Greece/aktogrammh/aktogrammh.shp")
# aktog_34N <- st_transform(aktog, crs = st_crs(features))
# aktog_ppap <- st_crop(aktog_34N,e)
# aktog_34Nsm <- st_simplify(st_crop(aktog_34N,e), dTolerance = 150)

# OSM Greek data -ENCODING IS AN ISSUE
places <- st_read("data/Greece/places.shp", options = "ENCODING=UTF8")

# roads <- st_read("data/Greece/roads.shp")
# waterways <- st_read("data/Greece/waterways.shp")
# natural <- st_read("data/Greece/natural.shp")
# points <- st_read("data/Greece/points.shp")

waterways <- st_transform(waterways, crs = 32634)
water_ppap <- st_crop(waterways,elev)

plot(water_ppap$geometry)

# lakes
lakes <- st_read("data/Greece/lakes.shp")
lakes_utm <- st_transform(lakes, crs = 32634)

mapview(lakes_utm)
```
### Get urban areas
```{r urban}
library(mapview)
places %>% 
  filter(population>1000) %>% 
  slice(c(172,460,463,597)) %>% 
  st_transform(crs = 32634) %>% 
 # st_intersection(b4) %>% 
  mapview()

localtowns <- places %>% 
  filter(population>1000) %>% 
  slice(c(172,460,463,597)) %>% 
  st_transform(crs = 32634)
crs=st_crs(4326)
localtowns$name <- c("Korinthos", "Loutraki", "Lecheo", "Perachora")

# Archaeological locations within study area
# 
archsites <- data.frame(osm_id = c(10,11,12,13, 14), 
           name = c("Skaloma", "Asprokampos","Schinos","Aremada", "Selki") ,
           type = c("archaeological locale","archaeological locale","archaeological locale",
                    "archaeological locale","archaeological locale"),
             population = c(0,0,0,0,0),
           longitude=c(22.9107094,22.9656586,23.0238265,22.97,22.98),
        latitude=c(38.0197058,38.0611656,38.0401937,38.025,38.037)) %>% 
   st_as_sf(coords = c("longitude", "latitude"),crs = 4326)


mapview(ppap_area)+mapview(archsites)+mapview(heraion_sf)

# Create a point for Heraion
heraion <- data.frame(osm_id = 1, 
           name = "Heraion", 
           type = "archaeological site",
             population = 0,
           longitude=c(22.8528769),
        latitude=c(38.0281432))

houseA <- data.frame(osm_id = 2, 
           name = "House A I", 
           type = "archaeological site",
             population = 0,
           longitude=c(22.859269),
        latitude=c(38.028981)) %>% st_as_sf(coords = c("longitude", "latitude"),crs = 4326)

# Convert the data frame to an sf object             
heraion_sf <- st_as_sf(heraion, coords = c("longitude", "latitude"),crs = 4326)
mapview(localtowns) + mapview(heraion_sf) + mapview(houseA)
```

### Get elevation
```{r elev}
# Get elevation data from SRTM
elev_gr <- getData("SRTM", lon = 22.6, lat = 38)
plot(elev_gr)
# elev <- crop(projectRaster(elev_gr, crs = CRS("+proj=utm +zone=34 +datum=WGS84 +units=m +no_defs")),e)
plot(elev)
plot(border)

## Reclassify the elevation raster to get rid of negative values
mat <- cbind(-500, 0, 0)
elev_gr <- reclassify(elev_gr, rcl = mat)
elev <- reclassify(elev, rcl = mat)
hist(elev_gr); 
hist(elev); hist(el)
plot(elev_gr)
plot(el);plot(b4, add = TRUE)

## Disaggregate to make raster smoother for detail views
elev_ppap <- disaggregate(elev, 5, method = "bilinear")
plot(elev_ppap)
```
### Get border
```{r border}
b_gr <- getData("GADM", country = "GRC", level = 0)
plot(b_gr)
mapview(b_gr) + mapview(lakes)
```

### Transform everything to EPSG 32634 - SKIP
```{r project-utm}
# border
b_gr <- st_transform(st_as_sf(b_gr), crs = 32634)
plot(b_gr)
```

## Load PPAP data
```{r load-ppap}
# Load data and transform to UTM34N (FAIMS already is)
ppap_area <- st_read("data/PPAPstudyarea.shp") %>% 
  st_transform(crs = 32634)
features <- st_read("data/Features20200130.shp")
units <- st_read("data/AllTeams.shp")
# SUattributes <- read_csv2("E:/Perachora2020/documentation/ALLSurveyUnits_noannotation.csv")
units <- inner_join(units, SUattributes, by= c("SUID" = "SurveyUnitID")) # only those that have a polygon (all but day 1)
```


## Bounding boxes and other details

```{r bboxes}
plot(units$geometry)
plot(features$geometry)

# Extents for tmap (useless otherwise)
b1 <- c(st_bbox(features) [1:2] - 250, st_bbox(features)[3:4] + 500)
b2 <- c(st_bbox(features)[1] - 2000, st_bbox(features)[2] - 10000, st_bbox(features)[3:4] + 5000)
b3 <- c(st_bbox(features)[1] - 5000, st_bbox(features)[2] - 15000, st_bbox(features)[3:4] + 15000)
b4 <- c(st_bbox(ppap_area)[1] - 2000, st_bbox(ppap_area)[2] - 10000, st_bbox(ppap_area)[3:4] + 5000)

# Bounding boxes for other uses and hand-drawn extents
b <- st_make_grid(units, n=1)
bb <- st_make_grid(features, n=1)

bbb <- as(eee, "SpatialPolygons") # make polygon out of a grid
bbb <- st_as_sf(bbb)
st_crs(bbb) <- 32634
bbbb <- st_make_grid(ppap_area$geometry, n=1)


plot(elev); plot(ee, col = "darkgreen", add = TRUE); plot(eee, col ="red", add = TRUE); plot(bb, add =TRUE);

bbb <- st_as_sf(as(eee, "SpatialPolygons"))
st_crs(bbb) <- 32634
box <- st_as_sf(as(ee, "SpatialPolygons"))
st_crs(box) <- 32634

plot(elev, main = "2020 PPAP study area");   plot(places$geometry, add =TRUE);plot(eee, col ="darkgreen", add = TRUE); plot(bbb$geometry, add = TRUE); plot(lakes$geometry, col = "blue", add = TRUE)
#plot(bbbb$geometry, col = "purple", add = TRUE); 

head(places)
```
## Greek map
```{r}
plot(elev_ppap, main = "2020 PPAP study area");  plot(box$geometry, border = "red", add =TRUE); plot(places$geometry, add =TRUE);plot(eee, col ="darkgreen", add = TRUE); #plot(bbb$geometry, add = TRUE); 
```


## Tmap Overview Greece

```{r greece-tmap}
# b_gr_simple <- st_simplify(b_gr, dTolerance = 250)
greece <- tm_shape(b_gr_simplecropped)+
  tm_polygons(col = "lightgrey", 
              border.col = "lightgrey")+
  tm_shape(heraion_sf) +
  tm_dots(col = "red",
          size = 0.5,
          alpha = 0.5) 
greece

```
### print greece
```{r}
pdf("figures/greece.pdf")#, colormodel = "gray")
greece
dev.off()
```


## Context of Perachora - in Greece
```{r study-area-BW-1}

overview <- tm_shape(el)+
  tm_raster(palette = "Greys", 
            title = "Elevation", 
            style = "pretty")+
  tm_shape(aktog_34Nsm) +
  tm_lines(lwd = 1) +
  tm_shape(bbb) +
  tm_borders(col = "red",
             lwd = 2) +
  tm_shape(ppap_area) +
  tm_borders(col = "purple") +
  tm_scale_bar(breaks = c(0, 10, 20),
               position = c("LEFT", "bottom"),
               text.size = 1) +
  tm_compass(position = c("RIGHT", "bottom"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "A. Sobotkova, 2021") +
  tm_layout(main.title = "Perachora Study Area",
          #  bg.color = "lightblue",
            inner.margins = c(0, 0, 0, 0))
overview

# Print to file
# pdf("figures/Figure01.pdf", colormodel = "gray")
# overview
# dev.off()
```

## Smaller -more detailed -  overview of Perachora
```{r overview-detail}
overview2 <- tm_shape(elev, bbox = b4) +
  tm_raster(palette = "Greys", 
            title = "Elevation (m)", 
            style = "pretty") + 
  tm_shape(lakes) +
  tm_polygons(col = "lightblue",border.col = "lightblue",
              lwd = 1) +
  tm_text("name",
          size = 0.75,
          col = "blue",
          ymod = -0.75,
          xmod = 0.5) +
  tm_shape(b_gr) +
  tm_borders(lwd = 3) +
             #col = "darkgrey")+ 
  tm_shape(ppap_area)+
  tm_borders(col = "purple",
             lwd = 2) +
  # tm_shape(bbb) +
  # tm_borders(col = "red",
  #            lwd = 2) +
  tm_shape(localtowns) +
  tm_bubbles(size = 0.3,
             shape = 21,
             col = "lightgrey")+
  tm_text("name",
          size = 0.75,
           ymod = -0.75,
          xmod = 0.5) +
  tm_shape(heraion_sf) +
  tm_bubbles(size = 0.1,
             col = "black")+
  tm_text("name",
          size = 0.75,
          ymod = -0.6,
          xmod = 0.5) +
  tm_shape(archsites) +
tm_bubbles(size = 0.3,  # keep the sizes suited to screen here
             shape = 21,
             col = "lightgrey")+
  tm_text("name",
          size = 0.75,
           ymod = -0.75,
          xmod = 0.5) +
  tm_scale_bar(breaks = c(0, 2.5, 5),
               position = c("RIGHT", "bottom"),
               text.size = .75) +
  tm_compass(position = c("RIGHT", "top"),
             type = "rose", 
             size = 2)
 # tm_credits(text = "A. Sobotkova, 2021") +
  # tm_layout(main.title = "Perachora Study Area 2020",
  #         #  bg.color = "lightblue",
  #           inner.margins = c(0, 0, 0, 0))
overview2


```
### print the map
```{r print-overview}
pdf("figures/Figure01b.pdf", width = 5)
overview2
dev.off()
```
### Print PPAP study area with inset
```{r}
library(tmaptools)
library(grid)

# change back to the plotting mode
#tmap_mode("plot")

pdf("figures/Figure01.pdf", width = 5)
# PPAP
overview2

# print insets
print(greece, vp=viewport(x= 0.17, y= 0.21, width= 0.3, height= 0.3))

dev.off()
```

## Survey densities
```{r survey-densities-features results}

a <- tm_shape(units, bbox = st_bbox(units)) + 
 # tm_polygons(col = "ivory", border.col = "ivory3")+
  tm_polygons(col="cornsilk2", border.alpha = 0) +
  tm_symbols(size = "MostRecentlyComputedAncient",
             col = "darkblue",
             scale = 3,
             title.size = "Sherd count per unit") +
   tm_shape(border_ppap, bbox = st_bbox(units)) +
   tm_borders(col = "grey", 
              lwd = 2)+
  tm_shape(st_as_sf(contours))+
  tm_lines(col = "lightgrey")+
  # tm_text(col = NA,
  #      text = "level",
  #       size = 0.5,
  #       remove.overlap = TRUE,
  #       along.lines = TRUE,
  #       overwrite.lines = TRUE, 
  #      bg.color = "white")+
  tm_shape(lakes) +
  tm_polygons(col = "lightblue",border.col = "lightblue",
              lwd = 1) +
  # tm_compass(type = "arrow", position = c("right", "top")) +
  # tm_scale_bar(position = c("right", "top"), breaks = c(0,0.1,0.2)) + 
  # tm_credits("Sobotkova 2020", position = c("left", "bottom")) +
  # tm_layout(title = "Artefact count per survey unit")
  tm_layout(title = "A")
a
b <- tm_shape(units, bbox = st_bbox(units), unit = "m") + 
  tm_polygons(col="cornsilk2", border.alpha = 0)+
  tm_shape(features) + 
  tm_symbols(shape = 6, col="#666666", size = 0.2) +
   tm_shape(border_ppap, bbox = st_bbox(units)) +
   tm_borders(col = "grey", 
              lwd = 2)+
   tm_shape(contours)+
  tm_lines(col = "lightgrey")+
  tm_shape(lakes) +
  tm_polygons(col = "lightblue",border.col = "lightblue",
              lwd = 1) +
  tm_compass(type = "arrow", 
             position = c("left", "bottom")) +
  tm_scale_bar(position = c("left", "bottom"),
               breaks = c(0,100,200), 
               text.size = 0.75) + 
  #tm_credits("Sobotkova 2020", position = c("right", "bottom"), size = 0.8) +
  #tm_layout(title = "Registered standing features") 
  tm_layout(title = "B")

current.mode <- tmap_mode("plot")
tmap_arrange(a,b)
```

### Print survey densities
```{r print, eval = false}
?pdf()
  pdf("figures/Figure11a_bw.pdf",
      colormodel = "gray", 
      width= 10, height = 6.5)
  # Arrange the two tmap layouts above one another
  current.mode <- tmap_mode("plot")
  tmap_arrange(a,b)  # add c if you want to see densities
  tmap_mode(current.mode)
  
  dev.off()
```



## Daily Progress
In the next chunk we visualise the daily progress and activity areas
```{r progress-days}
library(lubridate)
units$Date <- as_date(units$createdAtGMT)

rbind(heraion_sf, houseA)
         
daily <- tm_shape(units, 
         bbox = st_bbox(units), 
         unit = "m") + 
  tm_polygons(col = "Date", 
               #border.alpha = 0,
               palette = "Accent",
               title = "Daily coverage",
               n = 5) +
  tm_shape(border_ppap, bbox = st_bbox(units)) +
  tm_borders(col = "grey", 
              lwd = 2)+
  tm_shape(rbind(heraion_sf, houseA)) +
  tm_bubbles(size = 0.01,
             col = "white")+
  tm_text("name",
          size = 1,
          ymod = -0.6,
          xmod = 0.5) + 
  tm_compass(type = "arrow", 
               position = c("right", "top")) +
  tm_scale_bar(position = c("right", "bottom"),
                breaks = c(0,100,200)) 
  # tm_credits("Sobotkova 2021",
  # position = c("left", "bottom")) +
  #tm_layout(title = "PPAP 2020")
           # bg.color = "lightblue")

tmap_arrange(daily, visibility)  
```


## Visibility
```{r survey-vis}
units$Visibility <- factor(units$Visibility, 
                           levels = c("0-20%",
                                      "20-40%",
                                      "40-60%",
                                      "60-80%",
                                      "80-100%"))
#units$Vis <- as.numeric(units$Visibility)
#units$Visibility <- as.character(units$Visibility)
visibility <- tm_shape(units, 
         bbox = st_bbox(units), 
         unit = "m") + 
   tm_polygons(col = "Visibility", 
               #border.alpha = 0,
               palette = "Spectral",
               title = "Ground visibility",
               n = 5) +
  # tm_shape(aktog_ppap) +
  # tm_lines(col = "blue")+
  tm_shape(border_ppap, bbox = st_bbox(units)) +
   tm_borders(col = "grey", 
              lwd = 2)+
   tm_shape(rbind(heraion_sf, houseA)) +
   tm_bubbles(size = 0.01,
             col = "white")+
   tm_text("name",
          size = 1,
          ymod = -0.6,
          xmod = 0.5) + tm_compass(type = "arrow", 
               position = c("right", "top")) +
   tm_scale_bar(position = c("right", "bottom"),
                breaks = c(0,100,200)) 
  # tm_credits("Sobotkova 2021",
  # position = c("left", "bottom")) +
  #tm_layout(title = "PPAP 2020")
           # bg.color = "lightblue")

```

## Logarithmic artefact quantities
```{r artefact-log}
# wrangle the data 
units <- units %>% 
  mutate(AncientDensityHa = 
           MostRecentlyComputedAncient / Area_ha,
         AncientDensityLog = case_when(
    MostRecentlyComputedAncient == 0 ~ 0,
    MostRecentlyComputedAncient > 0 ~ log10(MostRecentlyComputedAncient))) 

# plot it
density_log <- tm_shape(units, 
         bbox = st_bbox(units), 
         unit = "m") + 
   tm_polygons(col="AncientDensityLog", 
               border.alpha = 0,
               title = "Density of artefacts (log10) ") +
   tm_shape(border_ppap, 
            bbox = st_bbox(units)) +
   tm_borders(col = "grey", 
              lwd = 2)+
   tm_shape(features) + 
    tm_symbols(shape = 6, 
               col="#666666",
               size = 0.2) +
   tm_compass(type = "arrow", 
               position = c("right", "top")) +
   tm_scale_bar(position = c("right", "bottom"),
                breaks = c(0,100,200)) # + 
  # tm_credits("Sobotkova 2021",
  # position = c("left", "bottom")) +
  # tm_layout(title = "Density of ancient artefacts")

density_log
```

```{r results-features}
featuremap <- tm_shape(units, bbox = st_bbox(units), unit = "m") + 
  tm_polygons(col="cornsilk2", border.alpha = 0)+
  tm_shape(features) + 
  tm_symbols(shape = 6, col="#666666", size = 0.2) +
  tm_shape(border_ppap) +
  tm_borders(col = "grey", 
              lwd = 2)+
  tm_compass(type = "arrow", position = c("right", "bottom")) +
  tm_scale_bar(position = c("right", "bottom"), breaks = c(0,100,200), text.size = 0.75) + 
  #tm_credits("Sobotkova 2020", position = c("right", "bottom"), size = 0.8) +
  tm_layout(title = "Registered standing features") 
featuremap

```

```{r with-relief-and-contours}
# Downsize the elevation raster
plot(el)
el_sm <- crop(el, box)
hist(el_sm)
el_ssm <- crop(el, bbb)

# Create contours
contours <- rasterToContour(el_ssm)
plot(contours);plot(features$geometry, add = TRUE); plot(aktog_34N$geometry, add = TRUE)

# tm_shape(el_ssm)+
#   tm_raster(palette = "Greens",
#             title = "Elevation",
#             style = "jenks",
#             n = 12)+
 tm_shape(features, bbox = bbb) + 
  tm_symbols(shape = 6, 
             col="#666666",
             size = 0.2) + 
  tm_shape(contours)+
  tm_lines(col = "lightgrey")+
    tm_shape(border_ppap) +
  tm_borders(col = "grey", 
              lwd = 2) + 
  tm_shape(units, unit = "m") + 
  tm_polygons(col="cornsilk2", border.alpha = 0)+
  
  
  tm_compass(type = "arrow", position = c("right", "bottom")) +
  tm_scale_bar(position = c("right", "bottom"), breaks = c(0,1,2), text.size = 0.75) + 
  #tm_credits("Sobotkova 2020", position = c("right", "bottom"), size = 0.8) +
  tm_layout(title = "Registered standing features") 

```
```{r extract-raster-data}
library(raster)
units$elev <- raster::extract(elev, st_centroid(units))

summary(units$elev)
hist(el_ssm, 
     xlab = "Elevation (msl)", 
     ylab = "Count",
     breaks = 25,
     main = "Elevation profile of Perachora study area and survey units (pink)"); 
hist(units$elev, col = rgb(1,0,0.2, 0.6), add = TRUE)

SU_elev <- zonal(elev, rasterize(units, elev, class = "SUID"), fun = "mean")
SU_elev

hist(el_ssm, 
     xlab = "Elevation (msl)", 
     ylab = "Count",
     breaks = 25,
     main = "Elevation profile of Perachora study area and survey units (pink)"); 
hist(SU_elev, col = rgb(1,0,0.2, 0.6), add = TRUE)
```

## Adding in pottery

```{r pottery}
library(tidyverse)
library(stringr)

pots <- read_csv("data/Perachora-pottery.csv")


sherds <- pots %>% 
  select(`FABRIC LEVIGATION: count and weigh by group, list number of handles, bases, rims, and body shards under #. Coarse#`, 'Med#', 'Fine#') %>% rename(Coarse = `FABRIC LEVIGATION: count and weigh by group, list number of handles, bases, rims, and body shards under #. Coarse#`, Med ='Med#', Fine = 'Fine#') %>% 
  mutate(CoarseCount = gsub("[A-Za-z]", "", Coarse),
         MedCount = gsub("[A-Za-z]", "", Med),
         FineCount = gsub("[A-Za-z]", "", Fine)
         ) %>% 
  mutate(CoarseSum = sapply(strsplit(CoarseCount, ' , '), function(x) sum(as.numeric(x))),
         MedSum = sapply(strsplit(MedCount, ' , '), function(x) sum(as.numeric(x))),
         FineSum = sapply(strsplit(FineCount, ' , '), function(x) sum(as.numeric(x)))) %>% 
           rowwise() %>% 
           mutate(TotalSherds = sum(CoarseSum, MedSum,FineSum, na.rm = TRUE)) 

sherds %>% 
  select(FineSum,MedSum,CoarseSum,TotalSherds) %>% 
  janitor::adorn_totals('row')

# Total pickups
a <- pi*9
density100_m <- 221/a
density103_m <- 737/a
541/a
```

